A Component-Based Sensor Network for \\ Environmental Monitoring

### 2009-03-12 docs/icccn2009/paper.tex
Updated the Hardware subsection.
git-svn-id: svn://localhost/netbeams/trunk@363 559edde2-a4d4-44e0-96fc-4035e40d7396

NetBEAMS supports an environmental sensor such as the YSI Sonde,
introduced in Section \ref{SEC_BACKGROUND}. In order to access its
sensor readings, we collocate it with an off-the-shelf embedded
computing platform called Gumstix \cite{gumstix01} (far left in Figure
\ref{FIG_NETBEAMS}). The Gumstix is a single-board computers with 256
Mb of flash and 256 Mb of SDRAM, and running a 600 MHz TI OMAP 3503
processor. Additional features can be added to the motherboard with
expansion cards connected via on-board buses. The motherboards draw
less than 250 mA @4V at 400 MHz and less than 50 mA while idling,
waiting for input.  Gumstix runs Linux 2.6 with the BusyBox utilities,
and uses the OpenEmbedded build environment to provide a complete Linux
environment and a large range of Linux applications. We use the JamVM
\cite{jamvm01} implementation of Sun Microsystem's virtual machine in order to
run the DSP.

The Gumstix connects to the RS232 serial port of the YSI Sonde via via a
null-modem cable.

@@

The schema is derived from DSPMessage. As a payload it adds an
unbounded sequence of sondeDataType. Each instance of this type
captures one timestamped measure point. The YSI Sonde component can be
configured via special DSP messages that define properties such as the
sampling frequency. These property messages originate from the web
frontend that we will explain in a subsequent section.

>>>>

NetBEAMS supports an environmental sensor such as the YSI Sonde,
introduced in Section \ref{SEC_BACKGROUND}. In order to access its
sensor readings, we collocate it with an off-the-shelf embedded
computing platform called Gumstix \cite{gumstix01} (far left in Figure
\ref{FIG_NETBEAMS}).  The Gumstix is an ARM architecture and comes in various 
hardware configurations.  A typical Gumstix configuration consists 
of a motherboard and one or more expansion boards which connect to the 
motherboard via on-board buses.  The motherboards draw less than 250 mA @4V 
at 400 MHz and less than 50 mA while idling, waiting for input.

In our project, we use the Verdex Pro XM4 motherboard with 64MB of SDRAM, 16MB of Flash, and 
a 400MHz Marvell PXA270 with XScale processor.  We also use two expansion boards:
the Console-VX which comes with three RS232 serial ports and the Netpro-VX which
features one 10/100baseT ethernet port.  The Gumstix runs Linux 2.6 with the 
BusyBox utilities, and uses the OpenEmbedded cross-compile build environment to 
provide a complete Linux environment and a large range of Linux applications for
the ARM architecture.  We use the JamVM \cite{jamvm01} implementation of Sun 
Microsystem's virtual machine in order to run the DSP. The Gumstix communicates 
with the YSI Sonde via an RS232 serial port using a null-modem cable since both 
the Sonde and the Gumstix are DTE devices. The serial port configuration is:

\begin{table}
\caption{\label{tab:RS232_Config}RS232 Serial Port Configuration}
\centering
\begin{tabular}{l l}
\hline
Baud Rate&9600\\
Data Bit&8\\
Parity&None\\
Handshake&None\\
\hline
\end{tabular}
\end{table}

\begin{table}
\caption{\label{TAB_RS232_Pinout}RS232 Serial Pinout}
\centering
\begin{tabular} {|l|l|l|}
\hline
Wire Color	&Pin Description        &DB-9\\ 
\hline
Yellow 		&RS232 TX    &2\\
Orange		&RS232 RX    &3\\  
Green  		&Alarm       &----\\ 
Grey			&RTS         &----\\
Blue  		&CTS         &----\\
Red			&+ 12V DC    &9\\
Black			&GND         &5\\  
Purple		&SDI-12      &----\\ 
Bare			&Shield      &----\\
\hline
\end{tabular}
\end{table}

@@

The schema is derived from DSPMessage. As a payload it adds an
unbounded sequence of sondeDataType. Each instance of this type
captures one timestamped measure point.

The YSI Sonde component can be
configured via special DSP messages that define properties such as the
sampling frequency. These property messages originate from the web
frontend that we will explain in a subsequent section.
